---
title: "Peer Assessment 1 - Reproducible Research - Jerome GRANSAC"
output: html_document
---

```{r, echo=FALSE}
setwd("C:\\Users\\Jerome\\Documents\\R\\work\\5 - Reproducible Research\\Ass 1")
```
##NOTE:  
###For this project, we use the "dplyr" library.      
```{r dplyr, warning=FALSE, echo=FALSE, message=FALSE}
library(dplyr)
```

###Loading and preprocessing the data  
####1. Read Data into data frame df  
```{r load_data}
df<-read.csv("activity.csv")
head(df)
##summary(df)

```

####2. Clean df by removing NA values  
```{r clean_data, echo=FALSE}

##formatC(as.numeric(df$interval), width=4, flag="0")
```



###What is mean total number of steps taken per day?  
I remove the missing values in the dataset.  
```{r rm_NA, }
dfNONA<-na.omit(df)
head(dfNONA)
```

####1. Calculate the total number of steps taken per day  
I have already removed the NA values above in dfNONA data frame. This is this data frame I am using below.
```{r total_number_steps_per_day, warning=FALSE}
## here we exclude the interval col
##aggregate(dfNONA[, 1], list(dfNONA$date), sum)
df_StepsPerDay<-dfNONA %>% select(date,steps) %>% group_by(date) %>% summarise_each(funs(sum))
##df_StepsPerDay
```

  
####2. Make a histogram of the total number of steps taken each day  
```{r hist, echo=FALSE, fig.height=4}
hist(as.data.frame(df_StepsPerDay)[,2], main="Histogram of the total number of steps taken each day", xlab="Total number of steps taken each day", ylab="Frequency", col="red")
```

####3. Calculate and report the mean and median of the total number of steps taken per day  
Mean of the total number of steps taken per day is:
```{r mean_display, echo=FALSE}
s<-summary(as.data.frame(df_StepsPerDay)[,2])
s["Mean"]
```
Median of the total number of steps taken per day is:
```{r median_display, echo=FALSE}
s["Median"]
```

###What is the average daily activity pattern?  
####1. Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days    
```{r step_interval, warning=FALSE,fig.height=4, echo=FALSE}
df_AvgStepsPerInterval5<-dfNONA %>% select(interval,steps) %>% group_by(interval) %>% summarise_each(funs(mean))
names(df_AvgStepsPerInterval5)<-c("Interval", "Avg_Steps")
plot( Avg_Steps~Interval,df_AvgStepsPerInterval5, xact="n", type="l",ylab="AVG number of steps taken", xlab="5mn intervals", main="Average number of steps taken every 5mn, averaged across all days")

```

####2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
``` {r max_number_steps, echo=TRUE, results="hide"}
Max_number<-filter(df_AvgStepsPerInterval5, Avg_Steps == max(df_AvgStepsPerInterval5$Avg_Steps))
##Max_number<-df_AvgStepsPerInterval5[which.max(df_AvgStepsPerInterval5$Avg_Steps),]$Interval
```

####The 5-minute interval that contains the max average number of steps across all days in the dataset is:  **`r Max_number[1]`**


###Imputing missing values
```{r missing_values, echo=TRUE,results="hide"}
missing_values<-sum(is.na(df))
```
#### 1. The total number of missing values in the dataset is **`r missing_values`** rows  

####2. Devise a strategy for filling in all of the missing values in the dataset.
The strategy I choose for filing in all of the missing values is using the mean for that 5-minute interval.   
For this, I need to retrieve the mean steps in the dataset for each 5-mn interval from a function below coded. Then when there is a NA step for a perticular interval, I get the mean of the steps for this interval with this function.
```{r function}
getMeanStepsToNA<-function()
{
  df_Filled<-df
  ##parsing all NA rows in steps from original dataset
  for (i in 1:nrow(df)) 
  {
    if(is.na(df[i,]$steps))
    {
      df_Filled[i,]$steps <- df_AvgStepsPerInterval5[df_AvgStepsPerInterval5$Interval==df[i,]$interval, ]$Avg_Steps
    }
  }
  df_Filled
}
```


####Create a new dataset that is equal to the original dataset but with the missing data filled in
I use the function below: **getMeanStepsFrom** for each interval where avg step is NA
```{r getMeanStepsToNA, cache=TRUE}
head(df)
df_Filled<-getMeanStepsToNA()
head(df_Filled)
```